Determining the alternatives for scalar implicature

### 2016-04-28 Paper/Paper.Rmd
adding information about trials per participant for three exps

On each trial participants were presented with a target scalar item from five scales, selected from the entailment items used in @vantiel2014b. These were embedded in a sentence such as, "In a recent restaurant review someone said they thought the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" They were prompted to generate three unique alternatives.

@@

Experiment 2 was conducted to approximate literal listener semantics---$p_{L_0}(m \mid u)$ in Equation (3)---for the same five pairs of scalar items used in Experiment 1. We included three conditions: two alternatives ("entailment"), four alternatives, and five alternatives. The two alternatives condition makes a test of the hypothesis that the two members of the classic Horn (entailment) scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four and five alternative conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four and five alternative conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiment 2 is to test whether the set of alternatives queried during literal semantic elicitation impacts literal semantic judgments. If it does we should see differences in these judgments for shared items between conditions

@@

Conditions were run sequentially. In each condition we recruited 30 participants from AMT. In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.} In the four alternatives condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternatives condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))` participants.

@@

The two alternatives condition only included entailment pairs taken from @vantiel2014b. The four alternatives condition included entailment pairs plus the top two alternatives generated for each scalar family by participants in Experiment 1. The five alternatives condition included the four previous items plus one more neutral item chosen from alternatives generated in Experiment 1.

@@

# for plotting
speakers$condition <- ifelse(speakers$expt == "e8", "two alts",
                                 ifelse(speakers$expt == "e10", 
                                        "four alts", "five alts"))
ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = condition)) + 
  geom_line() + 
  # facet_grid(.~ scale) + 
  facet_wrap(~wordPlots) +
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Condition")) + 
  xlab("star rating") +
  ylab("proportion compatible ('yes' responses)") + 
  # ggtitle("Scalar item literal semantic compatibility\nwith star ratings") +
  # scale_colour_solarized() + 
  scale_colour_manual(values=c("#dc322f","#268bd2","#859900")) + 
  theme_bw() +
  theme(strip.text.x = element_text(size = 6),
        legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        legend.title = element_text(size = 10),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 6))
```

@@

Participants were presented with a one-sentence prompt such as "Someone said they thought the food was \_____." with a target scalar item in the blank. Participants were then asked to generate a star rating representing the rating they thought the reviewer likely gave. Each participant was presented with all scalar items in a random order. The experimental setup is shown in Figure \ref{fig:stimuli_exp1}, right.

@@

compare_indices3 <- which(peloquinFrank_3Alts$words %in% check_words)
for (exp in exps) {
  three_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_3Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices3,
              compareDataName = exp,
              alphas = seq(0, 20, by = 0.1))
  res_table <- rbind(res_table, exp = three_alt_alphas[1, ])
}
compare_indices4 <- which(peloquinFrank_4Alts$words %in% check_words)
for (exp in exps) {
  four_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_4Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices4,
              compareDataName = exp,
              alphas = seq(0, 10, by = 0.1))
  res_table <- rbind(res_table, exp = four_alt_alphas[1, ])
}
compare_indices5 <- which(peloquinFrank_5Alts$words %in% check_words)
for (exp in exps) {
  five_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_5Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices5,
              compareDataName = exp,
              alphas = seq(0, 10, by = 0.1))
  res_table <- rbind(res_table, exp = five_alt_alphas[1, ])
}
res_table$row_num <- seq(1, nrow(res_table))
res_table <- res_table %>% 
  mutate(exp = ifelse(row_num %% 2 == 0, "Empirical 4-alts", "Empirical 2-alts"),
         model = ifelse(row_num < 3, "2-alts",
                           ifelse(row_num < 5, "3-alts",
                                  ifelse(row_num < 7, "4-alts", "5-alts"))),
         result = paste0("", round(cor, 2), " (alpha = ", alpha, ")")) %>%
  select(model, result, exp) %>%
  spread(exp, result)

@@

```{r fiveAltsScatter, fig.pos = "t", fig.width=3.5, fig.height=2.25, fig.cap = "Judgments for entailment items from the four alternatives condition of Experiment 3 are plotted against model predictions using the five alternatives data from Experiment 2. Colors show the star rating for individual judgments."}
best_alpha <- 4.5
best_run <- plyr::ddply(.data = peloquinFrank_5Alts, .variables =c("scale"), .fun = rsa.runDf,
                        semanticsVarName = "speaker.p",
                        quantityVarName = "stars",
                        itemVarName = "words",
                        alpha = best_alpha) %>%
  filter(words %in% check_words)

>>>>

On each trial participants were presented with a target scalar item from five scales, selected from the entailment items used in @vantiel2014b. These were embedded in a sentence such as, "In a recent restaurant review someone said they thought the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" Participants saw each scalar item in random order. In total each participant saw ten target scalar items (each entailment pair from the five scales selected from @vantiel2014b ).

@@

Experiment 2 was conducted to approximate literal listener semantics---$p_{L_0}(m \mid u)$ in Equation (3)---for the same five pairs of scalar items used in Experiment 1. We included three conditions: two alternatives ("entailment"), four alternatives, and five alternatives. The two alternatives condition makes a test of the hypothesis that the two members of the classic Horn (entailment) scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four and five alternative conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four and five alternative conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiment 2 is to test whether the set of alternatives queried during literal semantic elicitation impacts literal semantic judgments. If it does we should see differences in these judgments for shared items between conditions.

@@

Conditions were run sequentially. In each condition we recruited 30 participants from AMT. In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents who may have been confused by the question wording. Training trials in later conditions included clearer language.} In the four alternatives condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternatives condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))` participants.

@@

The two alternatives condition only included entailment pairs taken from @vantiel2014b for a total of 5 trails for each participant. The four alternatives condition included entailment pairs plus the top two alternatives generated for each scalar family by participants in Experiment 1 (100 trials per participant). The five alternatives condition included the four previous items plus one more neutral item chosen from alternatives generated in Experiment 1 (125 trials per participant).

@@

# for plotting
speakers$condition <- ifelse(speakers$expt == "e8", "two alts",
                                 ifelse(speakers$expt == "e10", 
                                        "four alts", "five alts"))
ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = condition, lty = condition)) + 
  geom_line() + 
  # facet_grid(.~ scale) + 
  facet_wrap(~wordPlots) +
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Condition")) + 
  xlab("star rating") +
  ylab("proportion compatible ('yes' responses)") + 
  # ggtitle("Scalar item literal semantic compatibility\nwith star ratings") +
  # scale_colour_solarized() + 
  scale_colour_manual(values=c("#dc322f","#268bd2","#859900")) + 
  theme_bw() +
  theme(strip.text.x = element_text(size = 6),
        legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        legend.title = element_text(size = 10),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 6))
```

@@

Participants were presented with a one-sentence prompt such as "Someone said they thought the food was \_____." with a target scalar item in the blank. Participants were then asked to generate a star rating representing the rating they thought the reviewer likely gave. Each participant was presented with all scalar items in a random order. Participants in the two alternatives condition gave total of 10 pragmatic judgments. Participants in the four alternatives condition gave a total of 20 pragmatic judgments. The experimental setup is shown in Figure \ref{fig:stimuli_exp1}, right.

@@

compare_indices3 <- which(peloquinFrank_3Alts$words %in% check_words)
for (exp in exps) {
  three_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_3Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices3,
              compareDataName = exp,
              alphas = seq(0, 20, by = 0.1))
  res_table <- rbind(res_table, exp = three_alt_alphas[1, ])
}
compare_indices4 <- which(peloquinFrank_4Alts$words %in% check_words)
for (exp in exps) {
  four_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_4Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices4,
              compareDataName = exp,
              alphas = seq(0, 10, by = 0.1))
  res_table <- rbind(res_table, exp = four_alt_alphas[1, ])
}
compare_indices5 <- which(peloquinFrank_5Alts$words %in% check_words)
for (exp in exps) {
  five_alt_alphas <- rsa.tuneDepthAlpha(data = peloquinFrank_5Alts,
              semanticsVarName = "speaker.p",
              quantityVarName = "stars",
              itemVarName = "words",
              groupName = "scale",
              compareIndices = compare_indices5,
              compareDataName = exp,
              alphas = seq(0, 10, by = 0.1))
  res_table <- rbind(res_table, exp = five_alt_alphas[1, ])
}
res_table$row_num <- seq(1, nrow(res_table))
res_table <- res_table %>% 
  mutate(exp = ifelse(row_num %% 2 == 0, "Empirical 4-alts", "Empirical 2-alts"),
         model = ifelse(row_num < 3, "2-alts",
                           ifelse(row_num < 5, "3-alts",
                                  ifelse(row_num < 7, "4-alts", "5-alts"))),
         result = paste0("r = ", round(cor, 2), " (alpha = ", alpha, ")")) %>%
  select(model, result, exp) %>%
  spread(exp, result)

@@

```{r fiveAltsScatter, fig.pos = "t", fig.width=3.5, fig.height=2.25, fig.cap = "Judgments for entailment items from the four alternatives condition of Experiment 3 are plotted against model predictions using the five alternatives data from Experiment 2. Colors show the star rating for individual judgments."}

check_words <- c("good", "excellent", "liked", "loved", "memorable", "unforgettable", "palatable", "delicious", "some", "all")
check_indices <- which(peloquinFrank_5Alts$words %in% check_words)
best_alpha <- rrrsa::rsa.tuneDepthAlpha(peloquinFrank_5Alts, semanticsVarName = "speaker.p",
                          quantityVarName = "stars", groupName = "scale", itemVarName = "words",
                          alphas = seq(1, 5, by = 0.1),  compareDataName = "e11",
                          compareIndices =check_indices)[1, 'alpha']

best_run <- plyr::ddply(.data = peloquinFrank_5Alts, .variables =c("scale"), .fun = rsa.runDf,
                        semanticsVarName = "speaker.p",
                        quantityVarName = "stars",
                        itemVarName = "words",
                        alpha = best_alpha) %>%
  filter(words %in% check_words)
